---
title: "Non-Linear Mass Loss Analysis"
output: html_document
date: "2025-03-20"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries}
devtools::load_all("D:/Github_Repos/litterfitter")
library(here)
library(tidyverse)
library(brms)
```
# Introduction
## Parameters, quantities, and posterior predictions
- Parameters, such as k, are highly model specific. They may also suffer from non-identifiability/equifinality (do the bootstrap distributions/pposterior distributions play this out?)

- Model agnostic quantities such as mean/median residence time, half-life. 

- Posterior predictions of the models themselves can have useful properties. The key here is that drawing from fitted distributions of the parameters can provide relevant error measures to propagate into secondary analysis

## Models of mass loss

M is Mt/M0, proportion of mass remaining, as is conventional
All models here assume normally distributed errors. Other error distributions, such as the Beta and related mixture models, are beyond the scope of this paper.

### Linear, time constant models of mass loss
The key thing to note is that these models assume a time-invariant slope.
- True linear
$$
M = -mt
$$
This model assumes that mass loss is lost at a linear, non-proportionally declining rate *m*

- Negative exponential (classic)
$$
M = e^{-kt}
$$

 For the negative exponential this is interpreted as a time-invariant 'proportional' mass loss. 

### Non-linear, time varying models of mass loss
Cornwell and Weedon (2014) and Manzoni et al (2012) explain these models in greater detail. 
Many examples exist, but two offer a great deal of flexibility to adequately capture the general range of mass loss trajectories observed in most mass loss time series. 


- Discrete series
$$
M_t = \frac{(1 - r) k_1 e^{-k_2 t} - (k_2 - k_1 r) e^{-k_1 t}}{k_1 - k_2}
$$

- Discrete parallel (AKA, bi- or double- exponential))
$$
M_t = \alpha e^{-k_1 t} + (1 - \alpha) e^{-k_2 t}
$$
```{r}
predict_discpar <- function(time, a, k1, k2){
  a * exp(-k1 * time) + (1 - a) * exp(-k2 * time)
}

discpar_sims <- expand_grid(time = seq(0,1000, by = 1),
                            a = c(0.75,0.5,0.25),
                            k1 = c(0.0001,0.01),
                            k2 = c(0.01,0.1,0.5)
                       )|>
  mutate(mass = predict_discpar(time, a, k1, k2))|>
  mutate(across(c(a,k1,k2), as_factor))

ggplot(discpar_sims, aes(x = time, y = mass, color = a))+
  geom_line()+
  facet_wrap(k1~k2)
```
Should really do a decent job of understanding k2 here - I thought erroneously that this would always be smaller than k1 on a leading trajectory but I was wrong. Has important implications for how to interpret k2 vs k1 (which appears to be conventional as k), e.g., with 'correcting' curves for leaching and then estimating Arrhenius temperature sensitivity - is this valid?
- Note here that k2 is the labile fraction

- Weibull (aka Weibull residence)
$$
M_t = e^{-\left( \frac{t}{\beta} \right)^\alpha}
$$
Beta is essentially a rate parameter interpretable as the time it takes to reach exp(-1) mass. 
Alpha is a shape parameter, controlling the concavity/convexity of the trajectory shape. Alpha >1 indicates a lagged mass loss trajectory (and could more closely align with a non-log linear model), alpha < 1 indicates a leading mass loss trajectory. The negative exponential model is a special case of the Weibull when alpha = 1. 


# Fitting (non-linear) models of mass loss
We will present two methods for fitting non-linear (and linear) models to mass loss data: estimating parameters via optimization (XXX of non-linear least squares?) or through Bayesian analysis (MCMC). These methods are of course also suitable for linear functions that can easily be fit through simple OLS methods. We could also linearly interpolate data or use LOESS methods but we are interested in highlighting two methods that are commonly used in probabilistic frameworks where error can be reasonably estimated in the model fits and propagated into model-derived quantities and downstream analysis of model predictions or parameters.

We will also turn to an existing dataset for example data, the Detrital Nutrients dataset (XXX link, citation). [XXX Data set description]. In particular, we will use a subset of these data that focused on stream litter, and further focus on time series that have number of observations (litter bag retrievals) greater than the number of parameters our models. 

```{r load Detrital Nutrients}
litter_df <- read_csv("https://github.com/robbinscalebj/revisiting-k/raw/refs/heads/main/data/derived_data/tidied_detrital_nutrients_data.csv")|> # tidied litter time series
  mutate(mass_prop = Mass_per_remaining/100)|>
  group_by(cohort_id)|>
  mutate(series_length = n())|>
  ungroup()|>
  filter(series_length > 5)|>
  rename(mass = "mass_prop", time = "Meas_Day")


litter_df|>
  group_by(cohort_id)|>
  arrange(cohort_id, desc(mass))|>
  slice_tail(n = 1)|>
  ggplot(aes(x = mass))+geom_histogram(bins = 15)+xlab("Final Mass Proportion")

```

`r length(unique(litter_df$cohort_id))` time series have at least 5 observations/retrievals during decomposition, with the final masses for the time series usually <40% remaining.

Examples:
Series 436 - lagged trajectory, Weibull best by each model GOF/parsimony estimator
Series 289 - leading trajectory, Weibull barely best model but discrete parallel also very good 
Series 118 - Constant variance, k is very good here

```{r define example time series}
example_series <- litter_df|>
  filter(cohort_id %in% c(118,289,436))

ggplot(example_series, aes(x = time, y = mass))+
  geom_point()+
  facet_wrap(.~cohort_id)+
  scale_y_continuous(trans = "log")

```



```{r parameter value distributions in Detrital Nutrients}
# load litter fits
# These were fit according to Robbins et al. (2025)
litter_fits <- readRDS(here("DetNut_litter_fits.Rds"))|>
  filter(cohort_id %in% {litter_df|>distinct(cohort_id)|>pull(cohort_id)})

tidy_fits <- 
  litter_fits|>
  mutate(negexp_converge_code = map_dbl(negexp_fit, ~pluck(., "optimFit", "convergence")),
         negexp_k = map_dbl(negexp_fit, ~pluck(., "optimFit", "par", 1)),
         negexp_AICc = map_dbl(negexp_fit, ~pluck(., "fitAICc")),
         negexp_logLik = map_dbl(negexp_fit, ~pluck(., "logLik")),
         negexp_BIC = map_dbl(negexp_fit, ~pluck(., "fitBIC")),
         negexp_pred = map(negexp_fit, ~pluck(., "predicted"))
  )|>
  mutate(weibull_converge_code = map_dbl(weibull_fit, ~pluck(., "optimFit", "convergence")),
         weibull_beta = map_dbl(weibull_fit, ~pluck(., "optimFit", "par", 1)),
         weibull_alpha = map_dbl(weibull_fit, ~pluck(., "optimFit", "par", 2)),
         weibull_AICc = map_dbl(weibull_fit, ~pluck(., "fitAICc")),
         weibull_logLik = map_dbl(weibull_fit, ~pluck(., "logLik")),
         weibull_BIC = map_dbl(weibull_fit, ~pluck(., "fitBIC")),
         weibull_pred = map(weibull_fit, ~pluck(., "predicted"))
         )|>
  mutate(discpar_converge_code = map_dbl(discpar_fit, ~pluck(., "optimFit", "convergence")),
         discpar_a = map_dbl(discpar_fit, ~pluck(., "optimFit", "par", 1)),
         discpar_k1 = map_dbl(discpar_fit, ~pluck(., "optimFit", "par", 2)),
         discpar_k2 = map_dbl(discpar_fit, ~pluck(., "optimFit", "par", 3)),
         discpar_AICc = map_dbl(discpar_fit, ~pluck(., "fitAICc")),
         discpar_logLik = map_dbl(discpar_fit, ~pluck(., "logLik")),
         discpar_BIC = map_dbl(discpar_fit, ~pluck(., "fitBIC")),
         discpar_pred = map(discpar_fit, ~pluck(., "predicted"))
         )|>
  mutate(discser_converge_code = map_dbl(discser_fit, ~pluck(., "optimFit", "convergence")),
         discser_r = map_dbl(discser_fit, ~pluck(., "optimFit", "par", 1)),
         discser_k1 = map_dbl(discser_fit, ~pluck(., "optimFit", "par", 2)),
         discser_k2 = map_dbl(discser_fit, ~pluck(., "optimFit", "par", 3)),
         discser_AICc = map_dbl(discser_fit, ~pluck(., "fitAICc")),
         discser_logLik = map_dbl(discser_fit, ~pluck(., "logLik")),
         discser_BIC = map_dbl(discser_fit, ~pluck(., "fitBIC")),
         discser_pred = map(discser_fit, ~pluck(., "predicted"))
         )|>
  mutate(contqual_converge_code = map_dbl(contqual_fit, ~pluck(., "optimFit", "convergence")),
         contqual_b = map_dbl(contqual_fit, ~pluck(., "optimFit", "par", 1)),
         contqual_a = map_dbl(contqual_fit, ~pluck(., "optimFit", "par", 2)),
         contqual_AICc = map_dbl(contqual_fit, ~pluck(., "fitAICc")),
         contqual_logLik = map_dbl(contqual_fit, ~pluck(., "logLik")),
         contqual_BIC = map_dbl(contqual_fit, ~pluck(., "fitBIC")),
         contqual_pred = map(contqual_fit, ~pluck(., "predicted"))
         )|>
  mutate(negexplim_converge_code = map_dbl(negexplim_fit, ~pluck(., "optimFit", "convergence")),
         negexplim_k = map_dbl(negexplim_fit, ~pluck(., "optimFit", "par", 1)),
         negexplim_a = map_dbl(negexplim_fit, ~pluck(., "optimFit", "par", 2)),
         negexplim_b = map_dbl(negexplim_fit, ~pluck(., "optimFit", "par", 3)),
         negexplim_AICc = map_dbl(negexplim_fit, ~pluck(., "fitAICc")),
         negexplim_logLik = map_dbl(negexplim_fit, ~pluck(., "logLik")),
         negexplim_BIC = map_dbl(negexplim_fit, ~pluck(., "fitBIC")),
         negexplim_pred = map(negexplim_fit, ~pluck(., "predicted"))
  )|>
  select(-c(negexp_fit, weibull_fit, discser_fit, discpar_fit, contqual_fit, negexplim_fit))

mod_gof <- tidy_fits|>
  #reshape for usability
  select(cohort_id, contains(c("AIC", "AICc", "BIC", "logLik")))|>
  pivot_longer(cols = -c(cohort_id), names_to = c("model", "gof_metric"), names_sep = "_", values_to = "value")|>
  filter(model %in% c("negexp","discpar","weibull"))|>
  group_by(cohort_id, gof_metric)|>
  mutate(best_in_cohort = case_when(gof_metric == "logLik" & value == max(value) ~ 1, #select highest when logLik
                                    gof_metric != "logLik" & value == min(value) ~ 1, # select lowest when AIC/BIC
                                    .default = 0))

tidy_fits|>
  select(negexp_k, discpar_a, discpar_k1, discpar_k2, weibull_alpha, weibull_beta)|>
  pivot_longer(everything(), names_to = "parameter")|>
ggplot(aes(x = value))+
  geom_density()+
  facet_wrap(.~parameter, scales = "free")

tidy_fits|>
  select(discpar_a, discpar_k1, discpar_k2)|>
  pivot_longer(everything(), names_to = "parameter")|>
ggplot(aes(x = value))+
  geom_histogram()+
  facet_wrap(.~parameter, scales = "free")+
  scale_x_continuous(limits = c(0,2))

tidy_fits|>
  ggplot(aes(x = discpar_a, y = discpar_k1))+geom_point()+scale_y_log10()+geom_smooth(method = "lm")
tidy_fits|>
  ggplot(aes(x = discpar_a, y = discpar_k2))+geom_point()+scale_y_log10()+geom_smooth(method = "lm")

tidy_fits|>
  ggplot(aes(x = discpar_k1-discpar_k2))+geom_histogram()
```

## Likelihood optimization and the litterfitter package
We used the litterfitter package, explained extensively in Cornwell and Weedon (2014), for the first approach. The litterfitter package minimizes the sum of squares using the L-BFGS-B algorithm, which is equivalent to maximizing the likelihood  under the assumed normally-distributed errors and constant variance. Error around the estimated parameters is done by bootstrapping: replicates of the time series are created by sampling with replacement, and then re-fit with the optimization procedure. This creates distributions of parameter values that convey uncertainty in the parameter value or create new predictions from the implied predictions of each parameter set. 

-litterfitter defaults assume timescale of years, which is not generally appropriate for aquatic litter breakdown that tends to proceed much more quickly.

### Example
litterfitter offers a lot of flexibility in easily fitting models to mass loss data and exploring the output. For now, we will use  time series that are best fit by several different models and represent three broad examples of decay trajectories. 



```{r litterfitter fits to example time series}
lf_iters<-9999 #takes 13 minutes on 1 core

if(file.exists(here("data/lf_example_fits.Rds"))){
  lf_example_fits <- readRDS(here("data/lf_example_fits.Rds"))
  
}else{
  

lf_example_fits <- example_series|>
    select(cohort_id, time, mass)|>
  group_by(cohort_id)|>
  nest()|>ungroup()|>
  mutate(negexp_fit = map(data, ~fit_litter(time = .$time, mass.remaining = .$mass, 
                            model = "neg.exp", iters = lf_iters,
                            upper = c(2),
                            lower = c(1e-08))),
         discpar_fit = map(data, ~fit_litter(time = .$time, mass.remaining = .$mass, 
                            model = "discrete.parallel", iters = lf_iters,
                            upper = c(1, 2, 2), #A, K1, K2 
                            lower = c(0, 1e-12, 1e-17))),
         weibull_fit = map(data, ~fit_litter(time = .$time, mass.remaining = .$mass, 
                            model = "weibull", iters = lf_iters,
                    lower = c(0.1,0.001), #beta, then alpha 
                    upper = c(10000,100))))|>
  mutate(negexp_bs = map(negexp_fit, ~bootstrap_parameters(., nboot = lf_iters, 
                            upper = c(2),
                            lower = c(1e-08))|>
                           unclass()|>as_tibble(.name_repair = ~ c("k", "V2"))|>select(k)),
        discpar_bs = map(discpar_fit, ~bootstrap_parameters(., nboot = lf_iters, 
                            upper = c(1, 2, 2), #A, K1, K2 
                            lower = c(0, 1e-12, 1e-17))|>
          unclass()|>as_tibble(.name_repair = ~ c("A", "k1", "k2", "V4"))|>select(A, k1, k2)),
        weibull_bs = map(weibull_fit, ~bootstrap_parameters(., nboot = lf_iters, 
                            lower = c(0.1,0.001), #beta, then alpha 
                            upper = c(10000,100))|>
          unclass()|>as_tibble(.name_repair = ~ c("beta", "alpha", "V4"))|>select(alpha, beta)))

saveRDS(lf_example_fits, here("data/lf_example_fits.Rds"))
}
#ggplot(lf_examples_bs$weibull_bs[[2]], aes(x = beta))+geom_density()

```
The distributions in these bootstrapped parameters suggest there is some instability, which is not too surprising if we are bootstrapping a relatively small sample where some samples can have particularly high leverage on the overall fit. 

## Bayesian estimation and brms
A Bayesian approach to fitting the XXX time series will produce similar results as the optimization procedure with the analytical models, although somewhat slower particularly if parallelizing the bootstrapping, but the difference in software and approach (including post-hoc analyses) may be preferable to some users. A primary consideration for this approach is the incorporation of priors. Just as the optimization routine required upper and lower bounds on parameter values, prior knowledge on parameter values can be specified as a statistical distribution. Compared to noninformative priors (e.g., flat priors), weakly informative priors have many benefits including better fitting behavior while simply incorporating what we already believe about a process - using previous empirical data or understanding, weakly informative priors can make highly unlikely parameters highly unlikely (Lemoine Oikos 2019). Just as the optimization approach, broadly synthesized datasets can be used to construct probability distributions that reflect our beliefs about what parameters are realistic and which are basically impossible. For example, a half-life for tree-leaf litter of 1 day (k= XXX) is exceedingly unlikely (quickly leached compounds rarely appear to account for more than 30% mass loss, Taylor and Barlocher 1996), and a half-life of 10 years is also unlikely. Additionally, the use of the term 'half-life' even implies another prior belief - that mass decreases as a function of time (although microbial colonization and extremely minor leaching can rarely cause this result during early stages of decomposition). Thus, priors can be drawn from continuous positive distributions such as the Gamma and lognormal. 

### Analytical models estimated with MCMC


```{r priors}

priors_df <- bind_rows(
  rlnorm(n = 5000, meanlog = -6, sdlog = 2)|>as_tibble()|>
    mutate(parameter = "negexp_k", median = median(value)),
  rgamma(n = 5000, 0.3,0.07)|>as_tibble()|>
    mutate(parameter = "weibull_alpha", median = median(value)),
  rlnorm(n = 5000, meanlog = 4.5, sdlog = 1.1)|>as_tibble()|>
    mutate(parameter = "weibull_beta", median = median(value))
)

ggplot(priors_df, aes(x = value))+
  geom_density()+
  facet_wrap(.~parameter, scales = "free")+
  theme_bw()

priors_df|>
  filter(case_when(
    parameter == "negexp_k" ~ value < 0.1,
    parameter == "weibull_alpha" ~ value < 20,
    parameter == "weibull_beta" ~ value < 1000
  ))|>
ggplot(aes(x = value))+
  geom_density()+
  geom_vline(aes(xintercept = median))+
  facet_wrap(.~parameter, scales = "free")+
  theme_bw()+ggtitle("Filtered distributions to show dominant probability mass")
# neg exp priors
ggplot(data = rlnorm(n = 1000, meanlog = -6, sdlog = 2)|>as_tibble())+geom_density(aes(x = value))+scale_x_continuous(limits = c(0,0.5))
#k_priors <- prior(lognormal(-6,2), nlpar = "k", lb = 0)
```

```{r discpar priors}
# discrete parallel alpha
ggplot(data = runif(5000, 0,1)|>as_tibble())+geom_histogram(aes(x = value))+
  geom_vline(aes(xintercept = median(value)))

#if alpha were a leachable fraction, we might want to assume a prior like rbeta(n = 5000,8,1.5)

#discpar k1 - want mass centered around 0.25, but may need to assume we can get out to about 15
ggplot(data = rlnorm(n = 1000, meanlog = -1, sdlog = 10)|>as_tibble())+geom_density(aes(x = value))+scale_x_continuous(limits = c(0,5))
#discpar k2
ggplot(data = rlnorm(n = 1000, meanlog = -1, sdlog = 4)|>as_tibble())+geom_density(aes(x = value))+scale_x_continuous(limits = c(0,0.5))
```

```{r discser priors}


```

```{r weibull priors}
# Weibull alpha
ggplot(data = rgamma(n = 5000, 1,0.2)|>as_tibble())+geom_density(aes(x = value))+
  geom_vline(aes(xintercept = median(value)))+scale_x_continuous(breaks = c(1,10,50), limits = c(0,30))
#weibull beta
ggplot(data = rlnorm(n = 5000, meanlog = 5.5, sdlog = 1)|>as_tibble())+geom_density(aes(x = value))+scale_x_continuous(limits = c(0,1000))+geom_vline(aes(xintercept = median(value)))


# sWeibull alpha
ggplot(data = rgamma(n = 5000, 2,0.2)|>as_tibble())+geom_density(aes(x = value))+
  geom_vline(aes(xintercept = median(value)))+scale_x_continuous(breaks = c(1,10,50), limits = c(0,2))

#sweibull beta
ggplot(data = rlnorm(n = 5000, meanlog = 6, sdlog = 1)|>as_tibble())+geom_density(aes(x = value))+scale_x_continuous(limits = c(0,1000))+geom_vline(aes(xintercept = median(value)))

```

```{r set up priors and formulae}

k_bf <- bf(mass ~ exp(-k*time), k ~1, nl = TRUE)
k_priors <- prior(lognormal(-6,2), nlpar = "k", lb = 0)

## Weibull
weibull_bf <- bf(mass ~ exp(-(time/beta)^alpha), beta + alpha ~1, nl = TRUE)
weibull_priors <- prior(gamma(1,0.2), nlpar = "alpha", lb = 0)+
  prior(lognormal(5.5,1), nlpar = "beta", lb = 0)

## Discrete parallel
discpar_bf <- bf(mass ~ a*exp(-k1 * time) + (1 - a) * exp(-k2 * time),
                 a + k1 + k2 ~ 1, nl = TRUE)

discpar_priors <- prior(uniform(0,1), nlpar = "a", lb = 0, ub = 1)+
  prior(lognormal(-1,4), nlpar = "k1", lb = 0)+
  prior(lognormal(-1,4), nlpar = "k2", lb = 0)

#discpar a, k1, and k2 are probably quite correlated


```

```{r fit brms models}
# example series: 118,289,436


k_118_brm <- brm(formula = k_bf, prior = k_priors,
                 data = example_series|>filter(cohort_id == 118), 
                 family = "gaussian", 
                 iter = 4000, chains = 4, cores = 4, warmup = 2000, 
                 file = here("data/brms_objects/k_118_brm"))
k_289_brm <- brm(formula = k_bf, prior = k_priors,
                 data = example_series|>filter(cohort_id == 289), 
                 family = "gaussian", 
                 iter = 4000, chains = 4, cores = 4, warmup = 2000, 
                 file = here("data/brms_objects/k_289_brm"))
k_436_brm <- brm(formula = k_bf, prior = k_priors,
                 data = example_series|>filter(cohort_id == 436), 
                 family = "gaussian", 
                 iter = 4000, chains = 4, cores = 4, warmup = 2000, 
                 file = here("data/brms_objects/k_436_brm"))


weib_118_brm <- brm(formula = weibull_bf, prior = weibull_priors,
                 data = example_series|>filter(cohort_id == 118), 
                 family = "gaussian", 
                 iter = 5000, chains = 4, cores = 4, warmup = 2000,
                 control= list(adapt_delta = 0.99), 
                 file = here("data/brms_objects/weib_118_brm")) # model hard to converge here
weib_289_brm <- brm(formula = weibull_bf, prior = weibull_priors,
                 data = example_series|>filter(cohort_id == 289), 
                 family = "gaussian", 
                 iter = 5000, chains = 4, cores = 4, warmup = 2000, 
                 file = here("data/brms_objects/weib_289_brm")) # pairs plot here has some non-identifiability in  beta/alpha
weib_436_brm <- brm(formula = weibull_bf, prior = weibull_priors,
                 data = example_series|>filter(cohort_id == 436), 
                 family = "gaussian", 
                 iter = 5000, chains = 4, cores = 4, warmup = 2000, 
                 file = here("data/brms_objects/weib_436_brm"))


discpar_289_brm <- brm(formula = discpar_bf, prior = discpar_priors,
                 data = example_series|>filter(cohort_id == 289), 
                 family = "gaussian", 
                 iter = 6000, chains = 4, cores = 4, warmup = 2000,
                 control= list(adapt_delta = 0.99),
                 file = here("data/brms_objects/discpar_289_brm"))


```




### Gaussian processes and GAMs (estimated with MCMC)

- A strong prior could be placed on k in the GAMs - shouldn't allow these to be very curvy at all. 


# Use of Posterior predictions/simulation

$$
k_{\mathrm{app}}(t) = \frac{d \ln{[A(t)]}}{dt} {(Manzoni_\ et_\ al._\ 2012)}
$$
This equation implies something quite intuitive - we can get instantaneous breakdown rates (a first order decay constant) for any arbitrary function A(t) by taking the time derivative of the natural log of the function. This also follows from the negative exponential model's derivative (Lecerf book chapter), which is a constant value since its time derivative is linear on the log scale.

- Note that negexp and discrete parallel have analytical solutions for derivatives 
## Median residence/transit time, half-life
- Choice of model matters here: show error estimates with k vs weibull and biexponential parameters - think about both bias and variance in the error
- Also relevant to demonstrate that Weibull's beta is (probably) useful here as a more reliable default parameter than k, when the data go to 37% mass loss. But still there are cases where something other than Weibull is a better fit and we would do best to propagate error into a model agnostic decay time


- How to do this with bootstrapping/optim appropriately? Single-fit bootstrapping vs non-parametric bootstrap solution
Probably need to compare regression estimates from one vs another under different non-linearity situations, and against the Bayesian posterior simulation.

Optim approach 1: Non-parametric bootstrap
- Obtain B sets of parameters, each set created from data resampling

Optim approach 2: Single fit covariance matrix: Single fit bootstrap
- Fit model
- get asymptotic covariance matrix by inverting the Hessian (i.e., asymptotic information matrix) - see eqn 3 in Fletcher and Jowett 2022
- Note that the hessian returned by optim may be numerically unstable, try to implement with numDeriv::hessian() - testing gave the same answer between the two Hessians
- See 'test_new_litterfitter_functionality.R" for test workspace and load of litterfitter fork and 'internal.R' from litterfitter fork.

"The inverse of the negative hessian is an estimator of the asymptotic covariance matrix"
https://stats.stackexchange.com/questions/27033/in-r-given-an-output-from-optim-with-a-hessian-matrix-how-to-calculate-paramet


-simulate B parameter sets from the multivariate normal, plug in the parameter values and obtain intervals as the produced distributions

- This is basically what gavin simpson does to get his GAM derivatives

## Continuous fragmentation losses


```{r joint model approach}
# model both time series and ALSO the difference between the quantities

```

## Two stage covariate analysis

### Temperature
- Direct approach
- Meta-analysis
What happens when we use k vs k with uncertainty propagated into the analysis

### Nutrients


# Limitations
- Linear analysis is really simple. Just need to be careful about whether we could be doing better. 
- Why not just plug in covariates of alpha and beta? 
- Why not use GAMs? 
-- GAMs not out-of-the-box constrained to 1 at t = 0
- Why assume normally distributed error (talking about a process misspecification rather than a distributional misspecification?)
  -- If we specify an Intercept with uncertainty (either estimated or as a latent value) rather than use the functional known of M=1|t=0
  -- Interesting thought: Are we more uncertain about t0.75 than t0.5 just because of errors growing along the function? Does this imply something about needing a link function? 