---
title: "Notes on Serial Weibull"
output: html_document
date: "2025-04-25"
---

This document is notes on a potential serial leaching-Weibull model of decomposition. This model fits some data quite well. This doc is now deprecated

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
litter_df <- read_csv("https://github.com/robbinscalebj/revisiting-k/raw/refs/heads/main/data/derived_data/tidied_detrital_nutrients_data.csv")|> # tidied litter time series
  mutate(mass_prop = Mass_per_remaining/100)|>
  group_by(cohort_id)|>
  mutate(series_length = n())|>
  ungroup()|>
  filter(series_length > 5)|>
  rename(mass = "mass_prop", time = "Meas_Day")

```


# The Parallel Olsen-Weibull (POW)
Parallel Negative Exponential - Generalized Exponential


Assumption here is that leaching proceeds rapidly (heavily baked into priors) and is followed by biologically dominated decomposition which has a possible lag in mass loss compared to the leaching process due to decomposer growth and colonization.

```{r}
pnge_function <- function(time, r, k1, alpha, beta) {

  M <- r * exp(-k1 * time)  + (1 - r) * exp(-(time / beta)^alpha)

}

pnge_bf <- bf(
  mass ~ r * exp(-k1 * time) + (1 - r) * exp(-(time / beta)^alpha),
  r + k1 + alpha + beta ~ 1,
  nl = TRUE
)

pnge_priors <- prior(beta(8,1.5), nlpar = "r", lb = 0, ub = 1)+ # makes leachable mass >0.5 very unlikely
  prior(lognormal(0.2,0.1), nlpar = "k1", lb = 0)+ #fractional mass loss where all leaching done (5% leachable remaining) in 2-3 days, with most mass around  
  prior(gamma(1,0.2), nlpar = "alpha", lb = 0)+
  prior(lognormal(6,1), nlpar = "beta", lb = 0)

pnge2_priors <- prior(beta(8,1.5), nlpar = "r", lb = 0, ub = 1)+ # makes leachable mass >0.5 very unlikely
  prior(lognormal(0.001,0.5), nlpar = "k1", lb = 0)+  
  prior(gamma(1,0.2), nlpar = "alpha", lb = 0)+
  prior(lognormal(6,1), nlpar = "beta", lb = 0)


```

```{r prior predictive simulation}
pnge_tib <- tibble(
  draw = 1:ndraw_pp,
  r = rbeta(ndraw_pp, 1.5, 8),
  k1 = rlnorm(ndraw_pp, meanlog = 0.001, sdlog = 1),
  beta  = rlnorm(ndraw_pp, meanlog = 5, sdlog = 1.5),
  alpha = rgamma(ndraw_pp, shape = 1, rate = 0.2)
)|>
  expand_grid(time = 0:1000)|>
  mutate(y = pow_function(r = r, beta = beta, alpha = alpha, k1 = k1, time = time))|>
  group_by(draw)|>
  filter(y >0.1)|>
  mutate(model = "pow")
  #filter(alpha <1)

pnge_tib_sum <- pow_tib|>
  group_by(time)|>
  summarize(mean = mean(y), upper = quantile(y, 0.025), lower = quantile(y, 0.975))

ggplot()+
  geom_line(data=pnge_tib|>group_by(draw)|>slice_sample(n = 5), aes(x = time, y = y, group = draw), alpha = 0.05)+
  geom_line(data=pnge_tib_sum, aes(x = time, y = mean), alpha = 0.8)+
  geom_line(data=pnge_tib_sum, aes(x = time, y = upper), alpha = 0.8, linetype = "dashed")+
  geom_line(data=pnge_tib_sum, aes(x = time, y = lower), alpha = 0.8, linetype = "dashed")+
  theme_bw()

```

```{r}


pnge_436_brm <- brm(formula = pnge_bf, prior = pnge_priors,
                 data = litter_df|>filter(cohort_id == 436), 
                 family = "gaussian", 
                 iter = 6000, chains = 4, cores = 4, warmup = 2000,
                 control= list(adapt_delta = 0.99))

pnge_436_brm <- brm(formula = pnge_bf, prior = sweibull3_priors,
                 data = litter_df|>filter(cohort_id == 436), 
                 family = "gaussian", 
                 iter = 6000, chains = 4, cores = 4, warmup = 2000,
                 control= list(adapt_delta = 0.99))



conditional_effects(sweib2_436_brm, "time")

blah<-tibble(time = seq(0,250, by = 1))|>add_epred_draws(sweib3_436_brm,ndraws = 1000)|>
  tidybayes::mean_hdi()

#p1 <- 
ggplot(blah)+
  geom_line(aes(x = time, y = .epred))+
  geom_line(aes(x = time, y = .upper), linetype = "dotted")+
  geom_line(aes(x = time, y = .lower), linetype = "dotted")+
  geom_point(data = litter_df|>filter(cohort_id == 436), aes(x = time, y = mass))
  scale_x_continuous(limits = c(0,100))
```
