---
title: "Notes on Serial Weibull"
output: html_document
date: "2025-04-25"
---

This document is notes on a potential serial leaching-Weibull model of decomposition. This model fits some data quite well. The issue is that posterior prediction is difficult given the use of the fmax() function in the formula, which is apparently a stan function and not in R. Extracting the estimated coefficients and predicting with those is very simple, but getting uncertainty around those predictions is not super easy. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
litter_df <- read_csv("https://github.com/robbinscalebj/revisiting-k/raw/refs/heads/main/data/derived_data/tidied_detrital_nutrients_data.csv")|> # tidied litter time series
  mutate(mass_prop = Mass_per_remaining/100)|>
  group_by(cohort_id)|>
  mutate(series_length = n())|>
  ungroup()|>
  filter(series_length > 5)|>
  rename(mass = "mass_prop", time = "Meas_Day")

```
- Discrete series-Weibull
Assumption here is that leaching proceeds rapidly (heavily baked into priors) and is followed by biologically dominated decomposition which has a possible lag in mass loss compared to the leaching process due to decomposer growth and colonization.

$$

M = r \cdot e^{-k_1 \cdot t} + (1 - r) \cdot e^{-(\frac{t}{\beta})^{\alpha}} \\
r &\sim \mathrm{Beta}(8,\ 1.5) \\
k_1 &\sim \mathrm{LogNormal}(0.2,\ 0.1) \\
\alpha &\sim \mathrm{Gamma}(1,\ 0.2) \\
\beta &\sim \mathrm{LogNormal}(6,\ 1)

$$
Proof of concept
- COuld calculate weibulls from simulations in Robbins et al 2025: Revisiting k based on Lecerf's publication.
```{r}
r_true     <- 0.8   # Fraction nonleachable pool
k1_true    <- 0.5  # Leaching rate
beta_true  <- 120   # Weibull scale
alpha_true <- 2.5   # Weibull shape


serial_weibull <- function(time, r, k1, beta, alpha){
(1 - r) * exp(-k1 * time) +
       r * exp(-((pmax(0, time - log(1000)/k1) / beta)^alpha))
}

serial_leachable <- function(time, r, k1, beta, alpha){
  (1 - r) * exp(-k1 * time)
}

serial_nonleachable <- function(time, r, k1, beta, alpha){
  r * exp(-((pmax(0, time - log(0.05)/k1) / beta)^alpha))
}
# Simulate true values
disc_comb <- tibble(time = seq(0,200, by = 2))|>
  mutate(M_leachable = serial_leachable(time, r = r_true, k1 = k1_true, beta = beta_true, alpha = alpha_true),
         M_nonleachable = serial_nonleachable(time, r = r_true, k1 = k1_true, beta = beta_true, alpha = alpha_true),
         M = M_leachable + M_nonleachable)


# Plot
ggplot(disc_comb, aes(x = time)) +
  geom_point(aes(y = M), color = "black", size = 1.5, alpha = 0.6) +
  #geom_line(aes(y = M_true), color = "blue", linewidth = 1.2) +
  labs(
    title = "Simulated Serial Decomposition Model with Weibull",
    y = "Remaining Mass (M)",
    x = "Time"
  ) +
  theme_minimal()

```

```{r}
## Serial weibull

sweibull_bf <- bf(mass ~ (1 - r) * exp(-k1 * time) + r * exp(-((fmax(0, time - log(0.05)/k1) / beta)^alpha)), 
                  r + k1 + beta + alpha ~1, nl = TRUE)
sweibull_priors <- prior(beta(8,1.5), nlpar = "r", lb = 0, ub = 1)+ # makes leachable mass >0.5 very unlikely
  prior(lognormal(0.2,0.1), nlpar = "k1", lb = 0)+ #fractional mass loss where all leaching done (5% leachable remaining) in 2-3 days, with most mass around  
  prior(gamma(1,0.2), nlpar = "alpha", lb = 0)+
  prior(lognormal(6,1), nlpar = "beta", lb = 0)


```

```{r}

sweib_289_brm <- brm(formula = sweibull_bf, prior = sweibull_priors,
                 data = example_series|>filter(cohort_id == 289), 
                 family = "gaussian", 
                 iter = 6000, chains = 4, cores = 4, warmup = 2000,
                 control= list(adapt_delta = 0.99)) #beta here is very poorly estimated because time series too short, for one thing

sweib_436_brm <- brm(formula = sweibull_bf, prior = sweibull_priors,
                 data = litter_df|>filter(cohort_id == 436), 
                 family = "gaussian", 
                 iter = 6000, chains = 4, cores = 4, warmup = 2000,
                 control= list(adapt_delta = 0.99))

blah <- fixef(sweib_436_brm)|>as.data.frame()|>rownames_to_column()|>mutate(rowname = str_remove(rowname, "_Intercept"))|>select(rowname, Estimate)|>
  pivot_wider(names_from = rowname, values_from = Estimate)|>
  expand_grid(time = seq(0, max(example_series|>filter(cohort_id == 436)|>select(time)), by = 0.1))|>
  mutate(M_leachable = serial_leachable(time = time, r = r, k1 = k1, beta = beta, alpha = alpha),
         M_nonleachable = serial_nonleachable(time, r = r, k1 = k1, beta = beta, alpha = alpha),
         M = M_leachable + M_nonleachable)

# can we use the R-based packages to predict if we define fmax in R? 

fmax <- function(x, y){
  max(x, y)
}
ggplot()+
  geom_point(data = example_series|>filter(cohort_id == 436), aes(x = time, y = mass))+
  geom_line(data = blah, aes(x = time, y = M))
```


```{r}
SerWeibull <- function(time, r, k1, alpha, beta) {

  M <- r * exp(-k1 * time)  + (1 - r) * exp(-(time / beta)^alpha)

}

sweibull2_bf <- bf(
  mass ~ r * exp(-k1 * time) + (1 - r) * exp(-(time / beta)^alpha),
  r + k1 + alpha + beta ~ 1,
  nl = TRUE
)
sweibull2_priors <- prior(beta(8,1.5), nlpar = "r", lb = 0, ub = 1)+ # makes leachable mass >0.5 very unlikely
  prior(lognormal(0.2,0.1), nlpar = "k1", lb = 0)+ #fractional mass loss where all leaching done (5% leachable remaining) in 2-3 days, with most mass around  
  prior(gamma(1,0.2), nlpar = "alpha", lb = 0)+
  prior(lognormal(6,1), nlpar = "beta", lb = 0)


sweib2_436_brm <- brm(formula = sweibull2_bf, prior = sweibull2_priors,
                 data = litter_df|>filter(cohort_id == 436), 
                 family = "gaussian", 
                 iter = 6000, chains = 4, cores = 4, warmup = 2000,
                 control= list(adapt_delta = 0.99))

sweib2_436_brm

conditional_effects(sweib2_436_brm, "time")

blah<-tibble(time = seq(0,250, by = 1))|>add_epred_draws(sweib2_436_brm,ndraw = 10)|>
  tidybayes::mean_hdi()

p1 <- ggplot(blah)+
  geom_line(aes(x = time, y = .epred))+
  geom_line(aes(x = time, y = .upper), linetype = "dotted")+
  geom_line(aes(x = time, y = .lower), linetype = "dotted")+
  geom_point(data = litter_df|>filter(cohort_id == 436), aes(x = time, y = mass))
```