---
title: "Figures for Presentation"
output: html_document
date: "2026-01-11"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(yardstick)

mytheme <- theme_bw(base_size = 18)+theme(panel.grid = element_blank(), 
                                          axis.title = element_text(face = "bold"), axis.text = element_text(face = "bold"))
```

# Negative exponential model example

```{r}
ne_df <- tibble(
  time = c(0, 2, 7, 28, 42),
  k = rep(0.05, length(time)),
  mass = exp(-k*time)
)|>
  mutate(mass_e = ifelse(time !=0, mass + rnorm(n = n(), mean = 0, sd = 0.15), mass))

ne_df2 <- tibble(
  time = seq(0,100, by = 0.05),
  k = rep(0.05, length(time)),
  mass=exp(-k*time)
)

ne_df2pt <- tibble(
  time = seq(0,100, by = 0.05),
  k = rep(0.05, length(time)),
  mass=exp(-k*time)
)|>
  filter(mass == 1 | round(mass, digits = 3) == 0.4)


ggplot(ne_df)+
  #geom_point(aes(x = time, y = mass), size = 4)+
  geom_point(aes(x = time, y = mass_e), color = "blue", size = 4)+
  xlab("Days")+ylab("Mass remaining (proportion)")+
  mytheme

ggplot(ne_df)+
  #geom_point(aes(x = time, y = mass))+
  geom_point(aes(x = time, y = log(mass_e)), color = "blue", size = 4)+
  geom_smooth(aes(x = time, y = log(mass_e)),method = "lm", se = FALSE)+
  xlab("Days")+ylab("Log(Mass remaining proportion)")+
  mytheme

ggplot(ne_df2)+
  geom_smooth(aes(x = time, y = mass),method = "gam", se = FALSE)+
  xlab("Days")+ylab("Mass remaining (proportion)")+
  scale_y_continuous(breaks = c(0,0.25,0.5,0.75,1))+
  mytheme

ggplot()+
  geom_smooth(data = ne_df, aes(x = time, y = log(mass)),method = "lm", se = FALSE)+
  geom_point(data = ne_df2pt, aes(x = time, y = log(mass)), size = 5)+
  xlab("Days")+ylab("Log(Mass remaining proportion)")+
  #scale_y_continuous(breaks = c(0,0.25,0.5,0.75,1))+
  mytheme

```





# Theoretical arguments 
Lecerf modeling

```{r}
pcdr <- expand_grid(
  detrital_mass = seq(0, 5.2, by = 0.01),
  c = c(-1,-0.5,0,0.22,0.5,1)
)|>
  mutate(pcdr = 0.25*(5.2-detrital_mass)^c)

ggplot(pcdr|>filter(c == 0.22))+
  geom_line(aes(x = detrital_mass, y = pcdr), linewidth = 1.5)+
  xlab("Detrital Mass (g)")+ylab(expression(bold("Per capita decomposer rate (" * d^-1 * ")")))+
  mytheme+scale_x_reverse()#+
  #facet_wrap(.~c, scales = "free_y")
```
```{r}
bm <- expand_grid(
  detrital_mass = seq(0, 5.2, by = 0.01),
  Bmax = c(400)
)|>
  mutate(B_M = (4*Bmax/(5.2^2))*(5.2-detrital_mass)*detrital_mass)

ggplot(bm)+
  geom_line(aes(x = detrital_mass, y = B_M), linewidth = 1.5)+
  xlab("Detrital Mass (g)")+ylab("Decomposer Biomass (g)")+
  mytheme+scale_x_reverse()+
  geom_segment(aes(y = 403, yend = 400, x = 2.5, xend = 5.2),  linetype = "dashed")+
  annotate("text", x = 5, y = 380, label = "bold(B[max])", parse = TRUE)
  #facet_wrap(.~c, scales = "free_y")
```

```{r}
leach <- expand_grid(
  time = seq(0, 6, by = 0.01),
  k = c(0.6)
)|>
  mutate(M_L = 2.8*exp(-k*time))

ggplot(leach)+
  geom_line(aes(x = time, y = M_L), linewidth = 1.5)+
  ylab("Leachable mass (g)")+ xlab("Days")+
  mytheme

```

# prior predct sims
```{r negexp prior pred sim}
ndraw_pp <- 100

# negexp
negexp_tib <- tibble(
  draw = 1:ndraw_pp,
  k = rlnorm(n = ndraw_pp, meanlog = -5, sdlog = 1.5)
)|>
  expand_grid(time = 0:250)|>
  mutate(y = exp(-k*time))|>
  group_by(draw)|>
  filter(y >0.1)|>
  mutate(model = "negexp")

ggplot(negexp_tib, aes(x = time, y = y, group = draw))+
  geom_line(alpha = 0.1)+
  xlab("Days")+ylab("Mass remaining (proportion)")+
  mytheme

ggplot(negexp_tib, aes(x = time, y = y, group = draw))+
  geom_line(alpha = 0.1)+
  xlab("Days")+ylab("Mass remaining (proportion)")+
  scale_y_continuous(trans="log", breaks = c(0.1, 0.2,0.4, 0.8,1))+
  mytheme
```

Could really think of these discrete parallel priors in two ways:
1) Representing past evidence that might include nonidentifiable parameters - weakly informative priors
2) Representing a very fast leachable fraction (e.g., ~90% leachable mass lost in first 5 days) - highly informative priors
```{r discrete parallel prior sim}
# discrete parallel alpha
ggplot(data = runif(5000, 0,1)|>as_tibble())+geom_histogram(aes(x = value))+
  geom_vline(aes(xintercept = median(value)))

ggplot(data = rbeta(n = 5000,10,1.5)|>as_tibble())+geom_histogram(aes(x = value))+
  geom_vline(aes(xintercept = median(value)))


#discpar k1 - want mass centered around 0.25, but may need to assume we can get out to about 15
ggplot(data = rlnorm(n = 1000, meanlog = -1, sdlog = 5)|>as_tibble())+geom_density(aes(x = value))+scale_x_continuous(limits = c(0,5))
#discpar k2
ggplot(data = rlnorm(n = 1000, meanlog = -1, sdlog = 4)|>as_tibble())+geom_density(aes(x = value))+scale_x_continuous(limits = c(0,0.5))

discpar_tib <- tibble(
  draw = 1:ndraw_pp,
  k1  = rlnorm(ndraw_pp, meanlog = -2, sdlog = 4),
  k2 = rlnorm(ndraw_pp, meanlog = -3, sdlog = 4),
  a = rbeta(n = ndraw_pp,10,1.5)
)|>
  expand_grid(time = 0:250)|>
  mutate(y = a*exp(-k1 * time) + (1 - a) * exp(-k2 * time))|>
  group_by(draw)|>
  filter(y >0.1)|>
  mutate(model = "discrete_parallel")

ggplot(discpar_tib, aes(x = time, y = y, group = draw))+
  geom_line(alpha = 0.1)+
  xlab("Days")+ylab("Mass remaining (proportion)")+
  mytheme

ggplot(discpar_tib, aes(x = time, y = y, group = draw))+
  geom_line(alpha = 0.1)+
  xlab("Days")+ylab("Mass remaining (proportion)")+
  scale_y_continuous(trans="log", breaks = c(0.1, 0.2,0.4, 0.8,1))+
  mytheme


```

```{r discrete series prior sim}
# discrete series r
ggplot(data = rbeta(n = 5000,1,2.5)|>as_tibble())+geom_histogram(aes(x = value))+
  geom_vline(aes(xintercept = median(value)))

#discser k1 - want mass centered around 0.25, but may need to assume we can get out to about 15
ggplot(data = rlnorm(n = 1000, meanlog = -2.25, sdlog = 1.5)|>as_tibble())+geom_density(aes(x = value))#+scale_x_continuous(limits = c(0,5))

#discser k2
ggplot(data = rlnorm(n = 1000, meanlog = -4.5, sdlog = 2)|>as_tibble())+geom_density(aes(x = value))+scale_x_continuous(limits = c(0,0.5))


discser_tib <- tibble(
  draw = 1:ndraw_pp,
  k1  = rlnorm(ndraw_pp, meanlog = -2.25, sdlog = 1.5),
  k2 = rlnorm(ndraw_pp, meanlog = -5, sdlog = 2.5),
  r = rbeta(n = ndraw_pp,1,2.5) # Check this
)|>
  expand_grid(time = 0:1000)|>
  mutate(y = (((1 - r) * k1 * exp(-k2 * time)) - ((k2 - k1 * r) * exp(-k1 * time)))/(k1 - k2))|>
  group_by(draw)|>
  filter(y >0.1)|>
  mutate(model = "discrete_series")

ggplot(discser_tib, aes(x = time, y = y, group = draw))+
  geom_line(alpha = 0.1)+
  theme_bw()
```


```{r weibull prior sim}
# Weibull alpha
ggplot(data = rgamma(n = 5000, 1,0.2)|>as_tibble())+geom_density(aes(x = value))+
  geom_vline(aes(xintercept = median(value)))+scale_x_continuous(breaks = c(1,10,50), limits = c(0,30))
#weibull beta
ggplot(data = rlnorm(n = 5000, meanlog = 5.5, sdlog = 1)|>as_tibble())+geom_density(aes(x = value))+scale_x_continuous(limits = c(0,1000))+geom_vline(aes(xintercept = median(value)))


#weibull

weibull_tib <- tibble(
  draw = 1:ndraw_pp,
  beta  = rlnorm(ndraw_pp, meanlog = 5.5, sdlog = 1),
  alpha = rgamma(ndraw_pp, shape = 1, rate = 0.2)
)|>
  expand_grid(time = 0:250)|>
  mutate(y = exp(-(time/beta)^alpha))|>
  group_by(draw)|>
  filter(y >0.1)|>
  mutate(model = "weibull")

ggplot(weibull_tib, aes(x = time, y = y, group = draw))+
  geom_line(alpha = 0.1)+
  xlab("Days")+ylab("Mass remaining (proportion)")+
  mytheme

ggplot(weibull_tib, aes(x = time, y = y, group = draw))+
  geom_line(alpha = 0.1)+
  xlab("Days")+ylab("Mass remaining (proportion)")+
  scale_y_continuous(trans="log", breaks = c(0.1, 0.2,0.4, 0.8,1))+
  mytheme

```

```{r parallel negative exponential-generalized exponential prior sim}


# pnge alpha
ggplot(data = rgamma(n = 5000, shape = 1, rate = 0.2)|>as_tibble())+geom_density(aes(x = value))+scale_x_continuous(limits = c(0,1000))+geom_vline(aes(xintercept = median(value)))

# pnge beta
ggplot(data = rlnorm(n = 5000, meanlog = 5, sdlog = 1.5)|>as_tibble())+geom_density(aes(x = value))+scale_x_continuous(limits = c(0,1000))+geom_vline(aes(xintercept = median(value)))

# pnge r
ggplot(data = rbeta(n = 5000, 1.5,8)|>as_tibble())+geom_density(aes(x = value))+scale_x_continuous(limits = c(0,1000))+geom_vline(aes(xintercept = median(value)))

# pnge k
ggplot(data = rlnorm(n = 5000, meanlog = 0.2, sdlog = 0.1)|>as_tibble())+geom_density(aes(x = value))+scale_x_continuous(limits = c(0,2))+geom_vline(aes(xintercept = median(value)))


# prior pred sim
pnge_tib <- tibble(
  draw = 1:ndraw_pp,
  r = rbeta(ndraw_pp, 1.5, 8),
  k1 = rlnorm(ndraw_pp, meanlog = 0.2, sdlog = 0.1), # fairly strong prior, but data seem to dominate likelihood in limited testing
  beta  = rlnorm(ndraw_pp, meanlog = 5, sdlog = 1.5),
  alpha = rgamma(ndraw_pp, shape = 1, rate = 0.2)
)|>
  expand_grid(time = 0:1000)|>
  mutate(y = pnge_function(r = r, beta = beta, alpha = alpha, k1 = k1, time = time))|>
  group_by(draw)|>
  filter(y >0.1)|>
  mutate(model = "pnge")#|>
  #filter(alpha >1)

pnge_tib_sum <- pnge_tib|>
  group_by(time)|>
  summarize(mean = mean(y), upper = quantile(y, 0.025), lower = quantile(y, 0.975))

ggplot()+
  geom_line(data=pnge_tib, aes(x = time, y = y, group = draw), alpha = 0.1)+
  #geom_line(data=pnge_tib_sum, aes(x = time, y = mean), alpha = 0.8)+
  #geom_line(data=pnge_tib_sum, aes(x = time, y = upper), alpha = 0.8, linetype = "dashed")+
  #geom_line(data=pnge_tib_sum, aes(x = time, y = lower), alpha = 0.8, linetype = "dashed")+
  theme_bw()

```

```{r prior pred sim plot}

pp_sims_df <- bind_rows(negexp_tib, discpar_tib, discser_tib, weibull_tib, pnge_tib)|>
  mutate(model_name = case_when(model == "negexp" ~ "Negative Exponential",
                                model == "discrete_parallel" ~ "Discrete Parallel",
                                model == "discrete_series" ~ "Discrete Series",
                                model == "weibull" ~ "Weibull",
                                model == "pnge" ~ "Parallel Negative-Generalized Exponential"))|>
  mutate(model_name = fct_relevel(model_name, "Negative Exponential","Discrete Parallel","Weibull","Parallel Negative-Generalized Exponential"))|>
  filter(model != "discrete_series")

ggplot(pp_sims_df, aes(x = time, y = y, group = draw))+
  geom_line(alpha = 0.1)+
  scale_x_continuous(limits = c(0,500))+
  facet_wrap(.~model_name)+
  ylab("Proportional Mass Remaining")+
  xlab("Time (Days)")+
  theme_bw()

```

# Model comparisons
```{r}
litter_df <- read_csv("https://github.com/robbinscalebj/revisiting-k/raw/refs/heads/main/data/derived_data/tidied_detrital_nutrients_data.csv")|> # tidied litter time series
  mutate(mass_prop = Mass_per_remaining/100)|>
  group_by(cohort_id)|>
  mutate(series_length = n())|>
  ungroup()|>
  filter(series_length > 3)|>
  rename(mass = "mass_prop", time = "Meas_Day")

litter_fits_long <- readRDS(url("https://raw.githubusercontent.com/robbinscalebj/revisiting-k/main/data/derived_data/DetNut_litter_fits.rds"))|>
  filter(cohort_id %in% {litter_df|>distinct(cohort_id)|>pull(cohort_id)})

# These were fit similarr to litter_fits, but more recently - this is the discrete parallel negative exponential-generalized exponential
nege_fits <- readRDS(here("DetNut_nege_fits.rds"))|>
  filter(cohort_id %in% {litter_df|>distinct(cohort_id)|>pull(cohort_id)})

tidy_fits <- 
  litter_fits_long|>
  mutate(negexp_converge_code = map_dbl(negexp_fit, ~pluck(., "optimFit", "convergence")),
         negexp_k = map_dbl(negexp_fit, ~pluck(., "optimFit", "par", 1)),
         negexp_AICc = map_dbl(negexp_fit, ~pluck(., "fitAICc")),
         negexp_AIC = map_dbl(negexp_fit, ~pluck(., "fitAIC")),
         negexp_logLik = map_dbl(negexp_fit, ~pluck(., "logLik")),
         negexp_BIC = map_dbl(negexp_fit, ~pluck(., "fitBIC")),
         negexp_pred = map(negexp_fit, ~pluck(., "predicted"))
  )|>
  mutate(weibull_converge_code = map_dbl(weibull_fit, ~pluck(., "optimFit", "convergence")),
         weibull_beta = map_dbl(weibull_fit, ~pluck(., "optimFit", "par", 1)),
         weibull_alpha = map_dbl(weibull_fit, ~pluck(., "optimFit", "par", 2)),
         weibull_AICc = map_dbl(weibull_fit, ~pluck(., "fitAICc")),
         weibull_AIC = map_dbl(weibull_fit, ~pluck(., "fitAIC")),
         weibull_logLik = map_dbl(weibull_fit, ~pluck(., "logLik")),
         weibull_BIC = map_dbl(weibull_fit, ~pluck(., "fitBIC")),
         weibull_pred = map(weibull_fit, ~pluck(., "predicted"))
         )|>
  mutate(discpar_converge_code = map_dbl(discpar_fit, ~pluck(., "optimFit", "convergence")),
         discpar_a = map_dbl(discpar_fit, ~pluck(., "optimFit", "par", 1)),
         discpar_k1 = map_dbl(discpar_fit, ~pluck(., "optimFit", "par", 2)),
         discpar_k2 = map_dbl(discpar_fit, ~pluck(., "optimFit", "par", 3)),
         discpar_AICc = map_dbl(discpar_fit, ~pluck(., "fitAICc")),
         discpar_AIC = map_dbl(discpar_fit, ~pluck(., "fitAIC")),
         discpar_logLik = map_dbl(discpar_fit, ~pluck(., "logLik")),
         discpar_BIC = map_dbl(discpar_fit, ~pluck(., "fitBIC")),
         discpar_pred = map(discpar_fit, ~pluck(., "predicted")))
         )|>
  mutate(discser_converge_code = map_dbl(discser_fit, ~pluck(., "optimFit", "convergence")),
         discser_r = map_dbl(discser_fit, ~pluck(., "optimFit", "par", 1)),
         discser_k1 = map_dbl(discser_fit, ~pluck(., "optimFit", "par", 2)),
         discser_k2 = map_dbl(discser_fit, ~pluck(., "optimFit", "par", 3)),
         discser_AICc = map_dbl(discser_fit, ~pluck(., "fitAICc")),
         discser_AIC = map_dbl(discser_fit, ~pluck(., "fitAIC")),
         discser_logLik = map_dbl(discser_fit, ~pluck(., "logLik")),
         discser_BIC = map_dbl(discser_fit, ~pluck(., "fitBIC")),
         discser_pred = map(discser_fit, ~pluck(., "predicted"))
         )|>
  mutate(contqual_converge_code = map_dbl(contqual_fit, ~pluck(., "optimFit", "convergence")),
         contqual_b = map_dbl(contqual_fit, ~pluck(., "optimFit", "par", 1)),
         contqual_a = map_dbl(contqual_fit, ~pluck(., "optimFit", "par", 2)),
         contqual_AICc = map_dbl(contqual_fit, ~pluck(., "fitAICc")),
         contqual_AIC = map_dbl(contqual_fit, ~pluck(., "fitAIC")),
         contqual_logLik = map_dbl(contqual_fit, ~pluck(., "logLik")),
         contqual_BIC = map_dbl(contqual_fit, ~pluck(., "fitBIC")),
         contqual_pred = map(contqual_fit, ~pluck(., "predicted"))
         )|>
  mutate(negexplim_converge_code = map_dbl(negexplim_fit, ~pluck(., "optimFit", "convergence")),
         negexplim_k = map_dbl(negexplim_fit, ~pluck(., "optimFit", "par", 1)),
         negexplim_a = map_dbl(negexplim_fit, ~pluck(., "optimFit", "par", 2)),
         negexplim_b = map_dbl(negexplim_fit, ~pluck(., "optimFit", "par", 3)),
         negexplim_AICc = map_dbl(negexplim_fit, ~pluck(., "fitAICc")),
         negexplim_AIC = map_dbl(negexplim_fit, ~pluck(., "fitAIC")),
         negexplim_logLik = map_dbl(negexplim_fit, ~pluck(., "logLik")),
         negexplim_BIC = map_dbl(negexplim_fit, ~pluck(., "fitBIC")),
         negexplim_pred = map(negexplim_fit, ~pluck(., "predicted"))
  )#|> comment out to retain other time series
  #left_join(nege_fits|>
#  mutate(nege_converge_code = map_dbl(nege_fit, ~pluck(., "optimFit", "convergence")),
    #     nege_r = map_dbl(nege_fit, ~pluck(., "optimFit", "par", 1)),
    #     nege_k = map_dbl(nege_fit, ~pluck(., "optimFit", "par", 2)),
    #     nege_alpha = map_dbl(nege_fit, ~pluck(., "optimFit", "par", 3)),
    #     nege_beta = map_dbl(nege_fit, ~pluck(., "optimFit", "par", 4)),
    #     nege_AICc = map_dbl(nege_fit, ~pluck(., "fitAICc")),
    #     nege_AIC = map_dbl(nege_fit, ~pluck(., "fitAIC")),
    ##     nege_logLik = map_dbl(nege_fit, ~pluck(., "logLik")),
    #     nege_BIC = map_dbl(nege_fit, ~pluck(., "fitBIC")),
    #     nege_pred = map(nege_fit, ~pluck(., "predicted")))
  #)|>
  #select(-c(negexp_fit, weibull_fit, discser_fit, discpar_fit, contqual_fit, negexplim_fit, -nege_fit))

mod_gof <- tidy_fits|>
  #reshape for usability
  select(cohort_id, contains(c("AIC", "AICc", "BIC", "logLik")))|>
  pivot_longer(cols = -c(cohort_id), names_to = c("model", "gof_metric"), names_sep = "_", values_to = "value")|>
  filter(model %in% c("negexp","discpar","weibull"))|>
  group_by(cohort_id, gof_metric)|>
  mutate(best_in_cohort = case_when(gof_metric == "logLik" & value == max(value) ~ 1, #select highest when logLik
                                    gof_metric != "logLik" & value == min(value) ~ 1, # select lowest when AIC/BIC
                                    .default = 0))


fit_comparisons <- mod_gof|>
  filter(best_in_cohort==1)|>
  group_by(gof_metric, model)|>
  count()|>
  ungroup()|>group_by(gof_metric)|>mutate(per = n/sum(n))

n_time_series <- length(unique(tidy_fits$cohort_id))



```


```{r comparison plots}
fit_counts.p <- fit_comparisons|>
  filter(gof_metric != "logLik" & gof_metric!="AIC")|>
  mutate(model = fct_relevel(model, "weibull", "negexp", after = Inf),
         model = fct_recode(model,Weibull = "weibull", NegExp = "negexp", `Continuous Quality` = "contqual", `Discrete Parallel` = "discpar", 
                            `Discrete Series` = "discser", NegExpGenExp = "nege"))|>
  ggplot(aes(x = gof_metric, y = n, fill = model), alpha =  1)+
  geom_bar(stat = "identity")+
  scale_fill_jco()+
  scale_y_continuous(breaks = c(0,50,100, n_time_series))+
  guides(fill = guide_legend("Model"))+
  xlab("Model Comparison Metric")+
  ylab("Count")+
  theme_classic()+
  theme(axis.title = element_text(face = "bold", size = 16), axis.text = element_text(size = 14, color = "black"),
        legend.title = element_text(size = 14), legend.text = element_text(size = 12))

fit_counts.p

```

# model errors
```{r}
fit_preds <- tidy_fits|>
  mutate(series_length = map_dbl(negexp_pred, ~length(.)))|> # get length of time series (same across all fits so just picking one set of predictions produced by a fit)
  relocate(series_length)|>left_join(
    litter_df|>select(cohort_id, time, Mesh_Size_Category, mass)|>
      mutate(mesh_cat = case_when(Mesh_Size_Category == "coarse" | Mesh_Size_Category == "fine" | Mesh_Size_Category == "open" ~ Mesh_Size_Category,
                              Mesh_Size_Category < 1 ~ "fine",
                              Mesh_Size_Category >= 1 ~ "coarse"))|>
      group_by(cohort_id)|>
      mutate(terminal_mass_prop = last(mass),
             terminal_meas_day = last(time))|>
      summarize(across(everything(), list))
  )|>
  select(-contains("negexplim"))|>
  select(series_length, time, mass, cohort_id, ends_with("pred"))|>
  group_by(cohort_id)|>
  unnest(cols = c(time, mass, negexp_pred, weibull_pred, discpar_pred))|>
  ungroup()|>
  mutate(across(contains("pred"), ~.-mass, .names = "{.col}_error"))

fit_rmse <- fit_preds|>
  filter(time != 0)|>
  group_by(cohort_id,series_length)|>
  summarize(across(ends_with("pred"), ~yardstick::rmse_vec(truth = mass, estimate = .)))|>
  pivot_longer(cols = contains("pred"), values_to = "rmse", names_to = "model")|>
  mutate(model = str_remove(model, "_pred"))|>
  filter(model %in% c("negexp","discpar","weibull"))
  
summed_rmse <- fit_rmse|>
  filter(model == "weibull")
quantile(summed_rmse$rmse, probs = c(0,0.9,0.75,0.5,0.25,0.1,0))  
  
fit_preds|>
  filter(time != 0)|>
  pivot_longer(contains("error"))|>
  ggplot(aes(x = value))+
  geom_histogram()+
  facet_wrap(.~name)

fit_preds|>
  mutate(series_length_category = case_when(series_length >8 ~ "9+",
                                            .default = as.character(series_length)))|>
  filter(time != 0)|>
  pivot_longer(contains("error"))|>
  ggplot(aes(x = value,  color = name))+
  geom_density()+
  facet_wrap(.~as_factor(series_length_category))+
  scale_color_jco()+
  theme_bw()




rmse_dstr.p <- fit_rmse|>
  mutate(model = fct_relevel(model, "weibull", "negexp", after = Inf),
         model = fct_recode(model,Weibull = "weibull", NegExp = "negexp", NegExpGenExp = "nege", `Continuous Quality` = "contqual", `Discrete Parallel` = "discpar", 
                            `Discrete Series` = "discser"))|>
  ggplot(aes(x = rmse,  color = model))+
  geom_density(linewidth = 1)+
  scale_linetype_manual(values = c(5,2,3,4,1))+
 # facet_wrap(.~as_factor(series_length))+
  xlab("RMSE (Litter mass proportion)")+
  ylab("Density")+
  scale_color_jco(name = "Model")+
  guides(linetype = "none")+
  theme_classic()+
  theme(axis.title = element_text(face = "bold", size = 16), axis.text = element_text(size = 14, color = "black"),
        legend.title = element_text(size = 14), legend.text = element_text(size = 12))
  
rmse_dstr.p    

```